
```{r}
mvt = read.csv("mvt.csv", stringsAsFactors=FALSE)

```

```{r}
str(mvt)
```
```{r}
# Convert the Date variable to a format that R will recognize:
mvt$Date = strptime(mvt$Date, format="%m/%d/%y %H:%M")

# Extract the hour and the day of the week:
mvt$Weekday = weekdays(mvt$Date)
mvt$Hour = mvt$Date$hour
```

```{r}
# Create a simple line plot - need the total number of crimes on each day of the week. We can get this information by creating a table:
table(mvt$Weekday)
```

```{r}
# Save this table as a data frame:
WeekdayCounts = as.data.frame(table(mvt$Weekday))

```

```{r}
# Load the ggplot2 library:
library(ggplot2)

# Create our plot
ggplot(WeekdayCounts, aes(x=Var1, y=Freq)) + geom_line(aes(group=1))  

```

```{r}
WeekdayCounts$Var1 = factor(WeekdayCounts$Var1, ordered=TRUE, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday"))

# Try again:
ggplot(WeekdayCounts, aes(x=Var1, y=Freq)) + geom_line(aes(group=1))

# Change our x and y labels:
ggplot(WeekdayCounts, aes(x=Var1, y=Freq)) + geom_line(aes(group=1),linetype =2) + xlab("Day of the Week") + ylab("Total Motor Vehicle Thefts")

```

```{r}
statesMap = map_data("state")
```

```{r}
ggplot(statesMap, aes(x = long, y = lat, group = group)) + geom_polygon(fill = "white", color = "black") 

```

Load the data from edges.csv into a data frame called edges, and load the data from users.csv into a data frame called users.

How many Facebook users are there in our dataset?
```{r}
edges <- read.csv("edges.csv")
users <- read.csv("users.csv")
```

```{r}
292/59
```
Out of all the students who listed a school, what was the most common locale?
```{r}
table(users$locale, users$school)
```

Is it possible that either school A or B is an all-girls or all-boys school?
```{r}
table(users$gender, users$school)
```
create a graph g describing our social network
Our edges are undirected -- if A is a Facebook friend of B then B is a Facebook friend of A. Therefore, we set the directed parameter to FALSE.
```{r}
#install.packages("igraph")
library(igraph)
```
```{r}
g = graph.data.frame(edges, FALSE, users)
```

we want to plot our graph
```{r}
plot(g, vertex.size=5, vertex.label=NA)
```

How many users are friends with 10 or more other Facebook users in this network?
```{r}
sum(degree(g) >= 10)
```

In a network, it's often visually useful to draw attention to "important" nodes in the network. While this might mean different things in different contexts, in a social network we might consider a user with a large number of friends to be an important user. From the previous problem, we know this is the same as saying that nodes with a high degree are important users.

To visually draw attention to these nodes, we will change the size of the vertices so the vertices with high degrees are larger. To do this, we will change the "size" attribute of the vertices of our graph to be an increasing function of their degrees:
```{r}
V(g)$size = degree(g)/2+2
plot(g, vertex.label=NA)
```
What is the largest size we assigned to any node in our graph?
```{r}
max(V(g)$size)
min(V(g)$size)
```
Thus far, we have changed the "size" attributes of our vertices. However, we can also change the colors of vertices to capture additional information about the Facebook users we are depicting.

When changing the size of nodes, we first obtained the vertices of our graph with V(g) and then accessed the the size attribute with V(g)$size. To change the color, we will update the attribute V(g)$color.

To color the vertices based on the gender of the user, we will need access to that variable. When we created our graph g, we provided it with the data frame users, which had variables gender, school, and locale. These are now stored as attributes V(g)$gender, V(g)$school, and V(g)$locale.

We can update the colors by setting the color to black for all vertices, than setting it to red for the vertices with gender A and setting it to gray for the vertices with gender B:
```{r}
V(g)$color = "black"

V(g)$color[V(g)$gender == "A"] = "red"

V(g)$color[V(g)$gender == "B"] = "gray"
```
```{r}
plot(g, vertex.label = NA)
```
Now, color the vertices based on the locale of the user.

```{r}
V(g)$color = "black"

V(g)$color[V(g)$locale == "A"] = "red"

V(g)$color[V(g)$locale == "B"] = "green"
```

```{r}
plot(g, vertex.label = NA)
```

Download the dataset "tweets.csv", and load it into a data frame called "tweets" using the read.csv() function, remembering to use stringsAsFactors=FALSE when loading the data.
```{r}
tweets <- read.csv("tweets.csv", stringsAsFactors = FALSE)
```

1) Create a corpus using the Tweet variable

2) Convert the corpus to lowercase

3) Remove punctuation from the corpus

4) Remove all English-language stopwords

5) Build a document-term matrix out of the corpus

6) Convert the document-term matrix to a data frame called allTweets

How many unique words are there across all the documents?
```{r}
library(tm)
corpus = VCorpus(VectorSource(tweets$Tweet)) 
corpus = tm_map(corpus, content_transformer(tolower))
corpus = tm_map(corpus, removePunctuation)
corpus = tm_map(corpus, removeWords, stopwords("english"))
frequencies = DocumentTermMatrix(corpus)
```
```{r}
allTweets = as.data.frame(as.matrix(frequencies))
```

```{r}
str(allTweets)
```

```{r}
#install.packages("wordcloud")
library(wordcloud)
```
Use allTweets to build a word cloud.

```{r}
arg1 = colnames(allTweets)
arg2 = colSums(allTweets)
```

```{r}
wordcloud(arg1,arg2, scale=c(2, 0.25))
```

